Multiway Spherical Clustering via \\ Degree-Corrected Tensor Block Models

### 2023-01-24 tensor_clustering/IEEE/ieee_revision/response_s1.tex

\newpage 
\begin{center}
    \textbf{Point-by-point Response to Reviewer 2}
\end{center}

\textit{Comments to the Author: This manuscript studies clustering under degree-corrected tensor block models (dTBMs). The dTBM is a generalization of the degree-corrected stochastic block model (DCSBM) for matrices/networks. It is also a generalization of the (non-degree) tensor block model (TBM) which is a higher-order extension of the stochastic block model (SBM).}

\textit{Contributions: The identifiability for the uniqueness of clustering under dTBMs is studied.
Statistical and computational lower bounds for clustering under dTBMs are established.
A two-stage, angle-based algorithm is proposed to achieve exact clustering in polynomial time under mild conditions. 
The last contribution is the most important one. The algorithm proposed uses the ``global-to-local strategy" which is commonly seen in community detection literature. An angle-based clustering is developed to handle the degree heterogeneity of dTBMs. Then theoretical analysis shows that it provably achieves the exact clustering. I feel this is a solid contribution, making the manuscript worth publication.}

\textit{Nevertheless, I have a few concerns and comments.}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]

\item \textit{One concern I have is that this manuscript is very similar to one of the authors' previous paper \cite{han2020exact} on clustering under the non-degree tensor block model. They share overall the same spirit, taste, and structure. For example:
The sessions for the computational and statistical lower bounds in this manuscript are very similar to the corresponding sessions in \cite{han2020exact} and they have the same results.
\cite{han2020exact} also gives a two-stage algorithm using the ``global-to-local strategy" and also achieves the exact clustering. \cite{han2020exact} gives the same threshold for the exact clustering as in this manuscript.
I think this manuscript should point out its connection with \cite{han2020exact}. The current version mentions that the method of \cite{han2020exact} fails under dTBMs but I think \cite{han2020exact} deserves more discussion given the similarity.}

\textbf{Response:} Thank you for your suggestion. We intentionally make the paper structure similar to~\cite{han2020exact} in order for easier comparison with non-degree models. Our results (e.g. the stat-comp lower bounds, the threshold for exact clustering) are all different from~\cite{han2020exact}. The confusion may come from the choice of our exposition: we chose the same notations as in~\cite{han2020exact} but actually they represent different quantities.

We have added a new Section~\ref{sec:tbm}, ``Comparison with Non-degree Tensor Block Model'', to discuss the connection and difference between our dTBM work and TBM in \cite{han2020exact} from three aspects: signal notions, theoretical results, and algorithms. Please see our response to {\bf \# 1, Major Comments from Editor} for details.

Briefly, we separate the discussion from three aspects: signal notion, theoretical results, and algorithm. We develop an angle-based signal notion to address the degree heterogeneity in dTBM. Our theoretical results are different with, and not easy extension of, the result in \cite{han2020exact}. Our algorithm share the classical local-to-global idea, but our angle-based iteration is specifically designed for dTBM and requires new proof techniques.  
%Particularly, our results in Section~\ref{sec:limits} are different with the result in \cite{han2020exact}. Though the critical values look the same, different signal notions lead to different interpretations of the results. The different parameter spaces and identifiability conditions also require different procedures to obtain the critical values. Also, our two-stage algorithm is different than that in \cite{han2020exact}. The angle-based update is specially designed for our degree-corrected model and brings a big challenge to establish the algorithm guarantee with polar coordinate based analysis.

\item \textit{My next comment is about whether Session III B (computational lower bound) can be much simplified. My understanding is that the non-degree tensor block model is a special case of dTBMs, and hence lower bounds givens in \cite{han2020exact} also hold under dTBMs. As a result, I also feel the 2nd contribution this manuscript claim (ie, statistical and computational lower bounds) are relatively minor.}

\textbf{Response:} Thank you for your suggestion. The Session III B (computational lower bound) cannot be derived from earlier non-degree work~\citep{han2020exact}. The short answer is that, our results show the similar conclusion but {\bf under different conditions.} While the TBM impossibility~\citep{han2020exact} provides a necessary condition for our dTBM impossibility, we find that {\bf such a condition is loose} in our contexts. This has motivated to develop new techniques to derive sharp lower bounds.

Inspired by your comment, we have revised the exposition of our theory. See our response to \textbf{\# 1 Major Comments from Editor} for details.

We quote the newly-added texts here:
\begin{quote}
    ``...\emph{Theoretical results.} In both works, we study the phase transition of TBM and dTBM with respect to the Euclidean and angle-based SNRs. We briefly summarize the results in \cite{han2020exact} and compare with ours.

    \textit{Statistical critical value:}
    \begin{align}
        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \quad \Delta_{\text{ang}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery;} \\
        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \quad \Delta_{\text{Euc}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery}.
    \end{align}

     \textit{Computational critical value:}
    \begin{align}
        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \quad \Delta_{\text{ang}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{ computationally efficient;} \\
        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \quad \Delta_{\text{Euc}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{computationally efficient}.
    \end{align}

The above comparison reveals three major differences.

 %Our dTBM phase transition re-discovers the statistical-computational gaps in TBM. %This finding indicates the intrinsic behavior distinctions among higher-order, matrix, and vector problem may be generally exists in many higher-order problems.
First, none of our results in Section~\ref{sec:limits} are corollaries of \cite{han2020exact}. Both models show the similar conclusion {\bf but under different conditions}. While the TBM impossibility~\citep{han2020exact} provides a necessary condition for our dTBM impossibility, we find that such a condition is often loose. There exists a regime of $\tS$ in which TBM problems are computationally efficient but dTBM problems are statistically impossible; see Example~\ref{example:euc_alg}. This observation has motivated us to develop the new signal notion $\Delta^2_{\text{ang}}$ for sharp dTBM phase transition conditions.

 Second, to find the phase transition, we need to show both the impossibility {\bf and} achievability when SNR is below {\bf and} above the critical value, respectively. While the TBM impossibility can serve as a loose condition of our dTBM impossibility, more efforts are required to show the achievability. In particular, since TBM is a more restrictive model than dTBM, the achievability in \cite{han2020exact} does not imply the achievability of dTBM in a larger parameter space. The latter requires us to develop new MLE and polynomial algorithms for dTBM achievability.  %For impossibility, the minimax bounds in dTBM search over a larger range of estimators than TBM. Though the thresholds for TBM impossibility can serve as the loose lower bounds of the dTBM impossibility, more efforts are required to avoid the gap between impossibility and achievablity.

Third, from the perspective of proofs, we develop new tools to handle the extra degree heterogeneity. In our Theorem~\ref{thm:stats}, we consider the profile MLE by maximizing out the nuisance degree parameter, while TBM~\citep{han2020exact} considers the usual MLE without degree parameter. In our Theorem~\ref{thm:comp}, we construct a rank-2 tensor to relate HPC conjecture to $\Delta^2_{\text{ang}}$, while TBM~\citep{han2020exact} constructs a rank-1 tensor to relate HPC conjecture to $\Delta^2_{\text{Euc}}$. The asymptotic non-equivalence between $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$ renders our proof technically more involved...''
   \end{quote}

   \begin{quote}
   ``... \emph{Signal Notion.} We emphasize that the signal notions are different between the two models. In particular, the Euclidean-based signal notion in TBM~\cite{han2020exact} fails to accurately describe the phase transition in our dTBM due to the possible heterogeneity in degree $\mtheta$. To compare, we denote our angle-based signal notion in~\eqref{eq:gamma} and the Euclidean-based SNR in \cite{han2020exact} as $\Delta_{\text{ang}}^2$ and $\Delta_{\text{Euc}}^2$, respectively:
\begin{equation}
     \Delta_{\text{ang}}^2 =  2(1 - \max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}} ), \quad \Delta_{\text{Euc}}^2 = \min_{a \neq b \in [r]} \onormSize{}{\mS_{a:} - \mS_{b:}}^2.
\end{equation}
By Lemma~\ref{lem:norm_diff} in the Appendix II, we have 
\begin{equation}\label{eq:signalcompare}
     \Delta_{\text{ang}}^2  \max_{a \in [r]}\onormSize{}{\mS_{a:}}^2 \leq \Delta_{\text{Euc}}^2.
\end{equation}
The above inequality indicates the condition $\Delta_{\text{Euc}}^2 \leq p^{\gamma}$ is {\bf sufficient but not necessary} for $\Delta_{\text{ang}}^2 \leq p^{\gamma}$. In fact, if we were to use $\Delta_{\text{Euc}}^2$ for both models, then the phase transition of dTBM can be arbitrarily worse than that for TBM.

Here, we provide an example to illustrate the dramatical difference between TBM and dTBM with the same core tensor.

\begin{example}[Comparison with Euclidean-based signal notion] \label{example:euc_alg} Consider a biclustering model with $\mtheta=1$ and an order-2 core matrix 
\begin{equation}
    \mS = \begin{pmatrix} p^{(\gamma+1)/2 } + 2  & 2 p^{(\gamma+1)/2} + 4\\
    2 & 4
    \end{pmatrix},\quad \text{with}\ \gamma \leq -1.
\end{equation}
The core matrix $\mS$ lies in the parameter spaces of TBM and our dTBM. Here, the constraint $\gamma \leq -1$ is added to ensure the bounded condition of $\mS$ in our parameter space in \eqref{eq:family}. The angle-based and Euclidean-based signal levels of $\mS$ are 
\begin{equation}
    \Delta_{\text{ang }}^2(\mS) = 0 \ \left(\leq p^{\gamma}\right), \quad \Delta_{\text{Euc}}^2(\mS) = 5 p^{\gamma + 1} \ \left(\geq p^{\gamma}\right).
\end{equation}
We conclude that TBM with $\mS$ achieves exact recovery with a polynomial-time algorithm; see \citet[Theorem 4]{han2020exact}. By contrast, the dTBM with the same $\mS$ and input $r=2$ violets the identifiability condition, and thus fails to be solved by all estimators; see our Theorem~\ref{thm:unique}.''
\end{example}

   \end{quote}
   We have revised our Section~\ref{sec:limits} to better present the lower bounds. See our response to \textbf{\# 1 Major Comments from Editor}.

\item \textit{My third comment is on Condition (7), i.e., ``balance" of $\theta$. This is quite artificial and impractical and is needed only for technical reasons in the theoretical analysis of the proposed algorithm. In fact, as far as I know, this (or similar condition) is not needed in DCSBM literature for community detection. I suspect (7) can be removed by some better proof technique or modification of the proposed method.}

\textbf{Response:} Thank you for your suggestion. See our response to \textbf{\#3(a) and 3(c), Reviewer 1} (page 7 of this response letter) for the detailed response.

While we agree the assumption is more for technical reasons, we suspect some ``balanced'' assumption or its variation seems necessary (and common in literature) for the achievability of computational critical values. 
Our constraints in $\tP$ are mild compared with previous literature. We have also added following comparison in Section~\ref{subsec:identify}:
\begin{quote}
    \begin{table}[h]
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c|c ccc}
    \hline
    Assumptions in parameter space& \cite{gao2018community}& \cite{han2020exact}& \cite{ke2019community} &Ours\\
    \hline
         Balanced community size &$\surd$ &$\surd$&$\surd$ & $\surd$  \\
        Balanced degree assumption & $\surd$&-&$\surd$ &$\surd$\\
         Bounded core tensor assumption & $\surd$ & $\times$ & $\surd$ &$\surd$\\
         Flexible in-group connections &$\times$ &$\surd$&$\surd$& $\surd$\\
         Gap among cluster centers & In-between cluster difference & Euclidean gap & Eigen gap & Angle gap\\
         \hline
    \end{tabular}
    }
    \caption{Parameter space comparison between previous work with our assumption.}
\end{table}

%We have added a new preliminary Section~\ref{sec:prelim} to discuss the regime with balanced $\mtheta$. The balance assumption is proposed to avoid the arbitrary bad phase transition of dTBM with respect to our signal notion. Similar regulations on degree also implicitly occur in DCSBM literature \cite{gao2018community, ke2019community}, in which more stringent constraints are applied on core tensor than ours. See our response to \textbf{ \# 3, Reviewer 1} for detailed discussion on balance assumption.

\end{enumerate}
\end{enumerate}

\newpage
\begin{center}
    \textbf{Point-by-point Response to Reviewer 3}
\end{center}

\textit{In general, this paper is well written and organized, numeric studies including simulation and real data application illustrate that the proposed model and method do have some merits. But I have several concerns regarding the theory and technical contribution.}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item \textit{Firstly, the mathematical formulation lacks motivation. From my knowledge, most existing degree-corrected block models are proposed for network data and Bernoulli model. While the proposed paper can be applied to Bernoulli as it could be taken as a special sub-Gaussian but I believe such treatment would not be suitable and/or optimal if one only cares about Bernoulli setting. Also, I notice several formulation inconsistencies. {For example, the author said they focus on symmetric tensor, but they assume the noise tensor $\mathcal{E}$ consists of independent entries, which does not make sense.}}

\textbf{Response:} Thank you for your suggestions. There seem to be three questions in this comment; we address them one-by-one.

\begin{enumerate}
\item Regarding ``motivation of general sub-Gaussian model''.
Our proposal of sub-Gaussian dTBM is intended to cover multiple types of data in applications. Continuous tensors arise commonly in fields such as genetics and medical image studies. From theoretical perspective, our sub-Gaussian framework allows both binary and continuous observations. Conversely, binary-specific methods may not be suitable for general (continuous) sub-Gaussian data. %We have added a motivating example in Section~\ref{subsec:motiv} to emphasize the popularity of continuous data. 
%\begin{quote}
   % `` ...
    %\textit{Gaussian tensor mixture model:} The Gaussian mixture model is a common model in pattern recognition and image processing. We say an order-$K$ sample tensor $\tY \in \bbR^{p_1 \times \cdots \times p_K}$ follows a tensor Gaussian distribution with mean tensor $\tS \in \bbR^{p_1 \times \cdots \times p_K}$ if $\text{vec}(\tY)$ is a Gaussian vector with mean $\text{vec}(\tS)$. Tensor Gaussian mixture model assumes the tensor data $\tY_i$ for all $i \in [N]$ follow the mixture tensor Gaussian distribution with density $f(\tY_i) = \sum_{a = 1}^r \pi_a \phi_a(\tS_a), \quad i \in [N]$, where $r$ is the number of clusters, $\pi_a$ is the mixture weight, $\tS_a$ is the mean tensor, and $\phi_a$ denotes the tensor Gaussian density of the $a$-th cluster with mean $\tS_a$ for all $a \in [r]$.

%Let $\tY \in \bbR^{N \times p_1 \times \cdots \times p_K }$ denote the combined data with slices $\tY(i,:) = \tY_i$ and $\tS \in \bbR^{r \times p_1 \times \cdots \times p_K}$ denote the combined mean tensor with slices $\tS(a,:) = \tS_a$. The tensor Gaussian mixture model with prior probability $\pi_a$ is equal to a special dTBM model
%\begin{equation}
   % \bbE[\tY] = \tS \times_1 \mM \times_2 \mI_{p_1} \times_3 \cdots \times_{K+1} \mI_{p_K},
%\end{equation}
%where $\mI_p$ is the identity matrix of dimension $p$, $\mM \in \{0,1\}^{N\times r}$ is the membership matrix for $N$ samples such that $\pi_a = |\{ i \in [N]: \mM_{i,a} = 1\} |/N$ for all $a \in [r]$. Assume that $\tS$ has a community structure. The tensor Gaussian mixture further becomes 
%\begin{equation}
   % \bbE[\tY] = \tC \times_1 \mM \times_2 \mTheta_1 \mM_1 \times_3 \cdots \times_{K+1} \mTheta_K \mM_K,
%\end{equation}
%where $\{ \mTheta_k, \mM_k\}_{k=1}^K$ refers to the heterogeneity and membership on the mean of Gaussian tensors. "
%\end{quote}

% First, our sub-Gaussian dTBM framework is well-motivated from the real life applications with multiple types of data. The general framework unifies a wide range of higher-order clustering models and facilities a comprehensive analysis for distinct observations. In contrast, the performances of existing Bernoulli methods are not guaranteed with general sub-Gaussian data. We have added more motivating examples to show and emphasize the importance of model flexibility. 
% {
% \color{blue}
% \begin{example}[Higher-order degree-corrected Gaussian mixture model] Gaussian mixture model is widely applied in applications including pattern recognition, image processing, and machine learning. We say an order-$K$ sample tensor $\tX \in \bbR^{p_1 \times \cdots \times p_K}$ follows a tensor Gaussian distribution with mean tensor $\tS \in \bbR^{p_1 \times \cdots \times p_K}$, denoted as $\tX \sim \tN(\tS; \mI_{p_1}, \ldots, \mI_{p_K})$, if $\text{vec}(\tY) \sim \tN( \text{vec}(\tS), \mI_{p_1 \cdots p_K})$, where $\mI_p$ is the identity matrix of dimension $p$.

### 2023-01-24 tensor_clustering/IEEE/ieee_revision/response_s1.tex

% Higher-order degree-corrected Gaussian mixture model assumes the tensor data $\tY_i \in \bbR^{p_1 \times \cdots \times p_K}$ for all $i \in [N]$ follow the mixture tensor Gaussian distribution with density 
% \begin{equation} \label{eq:gaussian_mix}
%     f(\tY_i) = \sum_{a = 1}^r \pi_a \phi_a( \theta_i \tS_a; \mI_{p_1}, \ldots, \mI_{p_K}), \quad i \in [N],
% \end{equation}
% where $\theta_i$ is the heterogeneity of the $i$-th observation, $r$ is the number of clusters, $\pi_a$ is the mixture weight, $\tS_a$ is the mean tensor  and $\phi_a$ denotes the tensor Gaussian density of the $a$-th cluster, respectively, for all $r \in [a]$.

% Let $\tY \in \bbR^{N \times p_1 \times \cdots \times p_K }$ denote the combined data with slices $\tY(i,:) = \tY_i$. Let $\tS \in \bbR^{r \times p_1 \times \cdots \times p_K}$ denote the combined mean tensor with slices $\tS(a,:) = \tS_a$. Further, we equip the mean tensor with community structure 
% \begin{equation}
%     \tS = \tC \times_2 \mM_1 \times_3 \cdots \times_{K+1} \mM_K,
% \end{equation}
% where $\tC \in \bbR^{r \times r_1 \times \cdots \times r_K}$ is the core tensor, and $\mM_k \in \{0,1\}^{p_k \times r_k}$ for all $k \in [K]$ are membership matrices. Consider the dTBM model
% \begin{equation}\label{eq:gaussian_mix_tbm}
%     \bbE[\tY] = \tC \times_1  \mTheta \mM \times_2 \mM_1 \times_3 \cdots \times_{K+1} \mM_K,
% \end{equation}
% where $\mTheta = \text{diag}(\theta_1, \cdots ,\theta_N)$ is the heterogeneity vector, $\mM \in \{0,1\}^{N \times r}$ is the membership matrix for $N$ samples. The higher-order degree-corrected Gaussian mixture model~\eqref{eq:gaussian_mix} is equivalent to our dTBM model~\eqref{eq:gaussian_mix_tbm} with prior probability $\pi_a = |\{ i \in [N]: \mM_{i,a} = 1\} |/N$ for all $a \in [r]$.
% \end{example}
% }
\item Regarding ``sub-optimality of Bernoulli as a special sub-Gaussian model''. We discuss below which of our results are still optimal for Bernoulli data, and which are not.

\begin{itemize}
\item {\bf sub-Algorithm I.} Our initialization gaurantee is not suitable for Bernoulli data. To address this, we have proposed a remedy for Bernoulli initialization with square unfolding; see Section~\ref{subsec:ber}. The Bernoulli initialization error rate $\tO(p^{-\floor{K/2}})$ is slightly weaker than Gaussian initialization error rate $\tO(p^{-K/2})$. Combining the Bernoulli initialization with sub-Algorithm II, our method achieves an exponential error rate for Bernoulli data.
%and beats the state-of-art~\citep{ke2019community}.

\item  {\bf sub-Algorithm II.} Our angle-based iteration guarantee is suitable for Bernoulli data. With Bernoulli data and a proper initialization, our iteration reduces the misclustering error from a polynomial rate to an exponential rate $\tO(\exp(-p^{K-1}))$, while the state-of-art algorithm for binary hypergraph \citep{ke2019community} leads to a polynomial error rate $\tO(p^{-2})$.

\item {\bf Theorems~\ref{thm:stats}-\ref{thm:comp}.} The phase transitions are not applicable for Bernoulli data. The analyses of statistical phase transition and computational impossibility are based on Gaussian noise, and the polynomial-time algorithm achievability requires a larger SNR of order $\tO(p^{-\floor{K/2}})$ for Bernoulli data. We have added a numerical experiment in Appendix I to verify the existence of statistical-computational gap in Bernoulli tensor clustering. 
\end{itemize}

New Section~\ref{subsec:ber}: 
\begin{quote}
    ``...

    \textit{Initialization guarantee for Bernoulli data.} The main difficulty to establish the initialization guarantee for Bernoulli observations lies in the denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}). We now provide a high-level explanation for the technical difficulty when applying Theorem~\ref{thm:initial} to Bernoulli observations.

    The derivation of Theorem~\ref{thm:initial} relies on the upper bound of the estimation error for the mean tensor in Lemma~\ref{lem:two-step_esterror}; i.e., with high probability
\begin{equation}\label{eq:prop1}
    \onormSize{}{\hat \tX - \tX}_F^2 \lesssim p^{K/2},
\end{equation}
where $\tX = \bbE\tY$ and $\hat \tX$ is defined in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}. Unfortunately, the inequality~\eqref{eq:prop1} holds only for i.i.d.\ sub-Gaussian observations, while Bernoulli observations are generally not identically distributed.

One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square} of Bernoulli tensor $\tY$. We summarize the procedure in sub-algorithm 3.

\begin{algorithm}[h!]
\caption*{\bf Sub-algorithm 3: Weighted higher-order initialization for Bernoulli observation}
\vspace{.15cm}
\begin{algorithmic}[1] 
\INPUT Bernoulli tensor $\tY \in \{0,1\}^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
\State  Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the nearly square unfolded tensor. Compute the estimate $\tX'$, where
\begin{equation}\label{eq:matrixsvd}
    \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
\end{equation}
\State Implement lines 3-5 of Sub-algorithm~\hyperref[alg:main]{1} with $\hat \tX$ replaced by $\hat \tX'$ in~\eqref{eq:matrixsvd}.
\OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
\end{algorithmic}
\end{algorithm}

\begin{prop}[Error for Bernoulli initialization]\label{cor:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ and Assumption~\ref{assmp:min_gap} holds. Assume that $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$. Let $ z^{(0)}$ denote the output of Sub-algorithm~3. With probability going to 1, we have
\begin{equation}\label{eq:ini_b}
 \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}, \quad \text{and} \quad L(z^{(0)}, z) \lesssim  {\sigma^2 r^K p^{-\floor{K/2}}}.
\end{equation}
\end{prop}

\begin{proof}[Proof of Proposition~\ref{cor:ber}] Following Lemma 7 in \cite{gao2018community}, with high probability, the new estimate satisfies
\begin{equation}
    \onormSize{}{\hat \tX' - \tX}_F^2 \lesssim p^{\ceil{K/2}}.
\end{equation} 
We obtain the upper bound by using $\hat \tX'$ in place of $\hat \tX$  in Theorem~\ref{thm:initial}.
\end{proof}

\begin{rmk}[Comparison with previous methods] The Bernoulli bound $\tO(p^{- \floor{K/2} })$ in Proposition~\ref{cor:ber} is relatively looser than the Gaussian bound \tO(p^{- K/2 })$ in Theorem~\ref{thm:initial}. The gap between Bernoulli and Gaussian error decreases as the order $K$ increases. Nevertheless, combining with angle iteration Sub-algorithm~\hyperref[alg:main]{2}, Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^{(K-1)}}$ at a price of a larger SNR.
%Nevertheless, the Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^ \floor{K/2}}$ with large SNR combining with our angle iteration Sub-algorithm~\hyperref[alg:main]{2}. 
The exponential rate is tighter than the rate error $\tO(p^{-2})$ in previous work~\citep{ke2019community}. The investigation of the gap between upper bound $p^{- \floor{K/2}}$ and the lower bound $p^{- K/2}$ for Bernoulli tensors will be left as future work. In numerical experiments, we will use our original initialization, Sub-algorithm 1, to verify the robustness to Bernoulli observations."
\end{rmk}

\end{quote}

New Appendix I: 
\begin{quote}
   `` \textbf{Bernoulli phase transition.} The first additional experiment verifies the statistical-computational gap in Section~\ref{sec:limits} under the Bernoulli model. Consider the Bernoulli model with $p = \{80, 100\}$, $r = 5$. We vary $\gamma $ in $ [-1.2, -0.4]$ and $[-2.1, -1.4]$ for matrix ($K=2$) and tensor $(K = 3)$ clustering, respectively. We  approximate MLE using an oracle estimator, i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from the true assignment. Figure~\ref{fig:phase_binary} shows a similar pattern as Figure~\ref{fig:phase}. The algorithm and oracle estimators have no gap in the matrix case, while an error gap emerges between the critical values $\gamma_{\text{stat}} = -2$ and $\gamma_{\text{comp}} = -1.5$ in the tensor case. Figure~\ref{fig:phase} suggests the statistical-computational gap in Bernoulli models."

\begin{figure}[htb]
    \centering
    \includegraphics[width = .8\columnwidth]{binary_phase.pdf}
    \caption{SNR phase transitions for Bernoulli dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.}
    \label{fig:phase_binary}
\end{figure}
\end{quote}

\item Regarding ``Formulation inconsistency". We agree that the original exposition is confusing and also unnecessary. We were attempting to simplify the notation by focusing on symmetric signal tensors (not necessarily symmetric data). Based on your comment, the revised paragraph in Section~\ref{sec:model} now reads

\begin{quote}
    ``... For ease of notation, we focus on the case with symmetric mean tensor $\bbE \tY$. This assumption simplifies the notation because all modes have the same $(\mTheta, \mM, z)$; the noise tensor $\tE$ and the data tensor $\tY$ are still possibly asymmetric. In general, we allow asymmetric mean tensor with $\{(\mTheta_k, \mM_k, z_k)\}_{k=1}^K$, one for each mode. The extension can be found in Appendix II...."
\end{quote}

In Section~\ref{sec:prelim}:
\begin{quote}
``...we provide...for general Gaussian dTBMs~\eqref{eq:model_margin} without symmetric assumptions. For general (asymmetric) Gaussian dTBMs, ....we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode. For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model. All results should be interpreted as the worst-case results across $K$ modes....''
\end{quote}

Consequently, we have revised our algorithm to reflect the asymmetry of data. We quote the revised algorithm in the last page of the response letter. We have also revised the notation in Theorems~\ref{thm:initial}-\ref{thm:refinement} by adding subscripts; that is, we now use $z_k^{(0)}$, $z_k^{(t+1)},\ldots$ in place of $z^{(0)}$, $z^{(t+1)}, \ldots$, throughout the paper.

%\begin{quote}
   % \begin{thm}[Error for weighted higher-order initialization] Consider the general sub-Gaussian dTBM with i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}. Assume $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$ and let $\Delta_{\mX}$ denote the minimal gap in mean tensor defined in~\eqref{eq:delta_x}.  Let $ z^{(0)}_k$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}. With probability going to 1, we have
%\begin{equation}
%    \ell(z^{(0)}_k, z) \lesssim {\sigma^2 r^K p^{-K/2}\over \Delta_{\mX}^2}.
%\end{equation}
%Further, assume $\mtheta$ is balanced as~\eqref{eq:degree}. We have
%\begin{equation}\label{eq:ini}
% \ell(z_k^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}} \quad \text{and} \quad L(z^{(0)}_k, z) \lesssim  {\sigma^2 r^K p^{-K/2}}
%\end{equation}
%\end{thm}

%\begin{thm}[Error for angle-based iteration] Consider the general sub-Gaussian dTBM with independent noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}. Let $\{z^{(0)}_k\}_{k=1}^K$ be the initialization for Sub-Algorithm~\hyperref[alg:main]{2} and $z^{(t)}_k$ be the $t$-th iteration output on $k$-th mode. Assume $\min_{i \in [p]}\theta(i) \geq c $ for some $c > 0$, the $\text{SNR} \geq \tilde C p^{-(K-1)}\log p$ for some sufficiently large constant $\tilde C$, and the initialization satisfies 
%\begin{equation}
%    L(z^{(0)}_k, z) \lesssim \frac{\Delta_{\min}^2}{r \log p}, \quad k \in [K].
%\end{equation}
%% With probability going to 1, there exists a contraction parameter $\rho \in (0,1)$ such that 
%\begin{align}\label{eq:final}
%    \ell(z, \hat z^{(t+1)}_k) \lesssim &\ \KeepStyleUnderBrace{
%   \text{SNR}^{-1}
%    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}_k). }_{\substack{\text{computational error}}}
%\end{align}
%\end{thm}

%\end{quote}

%Last, we have clarified the distribution of $\tE$ in equation~\eqref{eq:model_margin}. Throughout the paper, we consider the model with symmetric noise for simpler presentation; our theoretical results are based on the noise tensors with independent noise. We summarize the clarification here:
% \begin{quote}
%     ``..., and $\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a symmetric noise tensor consisting of independent zero-mean sub-Gaussian entries $(i_1,\ldots, i_K)$ for indices $i_1 \leq \cdots \leq i_K$ with variance bounded by $\sigma^2$. The entries in $\tE$ satisfy  $\tE(i_1,\ldots,i_K) = \tE(\pi(i_1),\ldots,\pi(i_K))$ for any permutation $\pi$ on $\{i_1, \ldots, i_K\}$. The symmetric constraint in $\tE$ is relaxed for general dTBM. ..."
% \end{quote}
\end{enumerate}

\item \textit{I also found some technical assumptions strong.  For example, in Eq. (3), the authors restricted the matricization of core tensor has norm bounded from below and above. I don't see why that is necessary. The author argues ``...so the core tensor $\tS$ is not trivially reduced to a lower rank", but the tensor block model should be identifiable even when $\tS$ has degenerated rank. This leads to another question on the defined SNR in Eq. (4) (which requires minimal $\tS$ sliced norm). It is doubtful that this is a true quantity to work on.}

\textbf{Response:} This comment concerns the lower and upper bounds for the core tensor. We address them separately.

\begin{enumerate}
\item Lower bound. Regarding ``...but tensor block model should be identifiable even when $\tS$ has degenerated rank.'' This is in fact what we intended to convey. Indeed, our remark of Theorem~\ref{thm:unique} reads: ``...the identifiability guarantee for the dTBM is stronger than classical Tucker model. .'' To avoid the confusion, we have revised the sentence into

\begin{quote}
``...Third, the lower bound constant $c_3>0$ requires no purely zero slide in $\tS$, and thus the model~\eqref{eq:model_tensor} is not trivially non-identifiable; see Example~\ref{example:c3}. 
...
\end{quote}

The lower boundness of core tensor slices is {\bf necessary} for the model identifiability. We have added following example in Section~\ref{subsec:identify} to illustrate the necessity of the lower bound $c_3$.

\begin{quote}
    `` ...

    \begin{example}[Non-identifiability with purely zero core slice] Consider an order-2 dTBM with core tensor $\mS = \begin{pmatrix} 0 & 0\\
    1 & -1
    \end{pmatrix}$ and mean tensor 
    \begin{equation}
        \tX =  \mTheta_1 \mM \mS   \mM^T \mTheta_2, \quad \text{with } \mM = \begin{bmatrix} 1 & 0 \\
        1 & 0 \\
        0 & 1 \\
        0 & 1 \\
        \end{bmatrix}, \quad \mTheta_1 =\mTheta_2= \text{diag}(1,1,1,1).
    \end{equation}
Replacing $\mTheta_1$ by $\mTheta'_1 = (3/2, 1/2, 1,1)$ leads to the same mean tensor $\tX$. "
\end{example}
\end{quote}
We note that the purely zero slice in $\tS$ always makes the degree heterogeneity indistinguishable. Therefore, we exclude such case for model identifiability.

\item Upper bound. The upper bound $c_4>0$ is a technical constraint to avoid unbounded entries in the core tensor; in practices, the constraint $\max_{a\in[r]} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large choice of $c_4 \geq \onorm{\tY}$. We have changed the wording in Section~\ref{subsec:identify} into

\begin{quote}
    ``... The upper bound $c_4>0$ is a technical constraint to avoid unbounded entries in the core tensor. We introduce this assumption for proof convenience; in practical algorithms, the constraint $\max_{a\in[r]} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large choice of $c_4 \geq \onormSize{}{\tY}_F$. "
\end{quote}
